---
title: "In-class 3"
author: "Lucas Vial"
date: "Sep 2 2024"
date-modified: "Date"
execute: 
  eval: false
  echo: true
  freeze: true
  warning: false
---

## Spatial Point Patterns Analysis 

### SpatStat

Installing maptools - Old version because we are using outdated software

```{r}
install.packages("maptools", repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

sf \> sp

Issue 2: Creating coastal outline

```{r}
mpsz_sf <- st_read(dsn = "../Hands-On/data/data_3/data/", 
                layer = "MP14_SUBZONE_WEB_PL")
childcare_sf <- st_read("../Hands-On/data/data_3/data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
sg_sf <- mpsz_sf %>%
  st_union()
```

List of different sub-packages for spat stat (looked at geom)

create a point pattern: ppp(x, y, owin)

as.ppp convert data to ppp (this one is more important)

as.owin convert data to owin (window or the squares that make up grids)

add maptools:: to the start of as.SpatitalGridDataFrame to fix it! -\> or you could just do it properly like I did

## Take home code example

acled_sf \<- read.csv() %\>%

st_as_sf(coords = c("longitude", "latitude"), crs=4326) %\>%

st_transform(crs = 32647) %\>%

mutate(event)data = dmy(event_date))
