---
title: "In-class 4"
author: "Lucas Vial"
date: "Sep 9 2024"
date-modified: "Date"
execute: 
  eval: true
  echo: true
  freeze: true
  warning: false
---

## Forest Fires in Sumatra, Indonesia 

### STKDE - Spatio-Temporal KDE (Kernel Density Estimator) 

KDE but you have one for distance in space and one in time

### Imports and Data Wrangling 

```{r}
pacman::p_load(sf, tmap, tidyverse, raster, spatstat, sparr, dplyr)
```

```{r}
kbb_sf <- st_read(dsn = "data/4data/rawdata", 
  layer = "Kepulauan_Bangka_Belitung") %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748) #This number is a EPSG 

```

EPSG - A number to refer to an area on the (world) map

```{r}
kbb_owin <- as.owin(kbb_sf)
```

```{r}
class(kbb_owin)
```

Make the csv (long, lat) into workable data :)

```{r}
fire_sf <- read_csv("data/4data/rawdata/forestfires.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"),
           crs = 4326) %>% 
  st_transform(crs = 32748)
```

```{r}
fire_sf <- fire_sf %>% 
  mutate(DayofYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date,
                           label = TRUE,
                           abbr = FALSE)) # January or JAN ??
```

\^\^ this code makes acq data numeric - makes them into new columns

### Plotting

**lubridate** is the timedate package used in this course (more talk about tidyverse 😻)

Be aware of Interactive mode being on or off

The Below code is very literal - Its drawing the polygon of kbb_sf and then adding the dots over the top

```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
tm_shape(fire_sf)+
  tm_dots()
```

Each Month Plot

```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
tm_shape(fire_sf)+
  tm_dots(size = 0.1)+
tm_facets(by="Month_fac",
          free.coords = FALSE, #will be weird and zoomed in
          drop.units = TRUE)
```

```{r}
# Doesn't Work - 
fire_month <- fire_sf %>%
  dplyr::select(Month_num)
```

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
```

### Computing and Plotting SPKDE

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

h = Spatial, which is 15102 = 15 km

lambda = Temporal, which is 0.0304 = 0.03 months

Notice how fig height and width makes the graphs looks half decent

```{r}
#| fig-width: 12
#| fig-height: 10
tims <- c(7,8,9,10,11,12)
par(mfcol=c(2,3))
for(i in tims){
  plot(st_kde, i,
       override.par=FALSE,
       fix.range=TRUE,
       main=paste("KDE at month",i))
}
```

Theres a way to make it into a gif where you can do it day by day

(he teased the code but didnt tell us how to do it 😠)
