---
title: "In-class 5"
author: "Lucas Vial"
date: "Sep 16 2024"
date-modified: "Date"
execute: 
  eval: true
  echo: true
  freeze: true
  warning: false
---

Arrived only 5 minutes late!

## Lectured Slide Notes

Spatial Weights (wij), How we define spatial neighbourhoods. (are these two things neighbours, wij is how we decide!)

-   Adjacency criterion: Is it right next to each other? Binary Output

-   Distance criterion: If its within a set a parameter of distance (use your little dots) - if location j is within distance d from i

-   You can also give them their own weights; not a pass fail its close to the one

Contiguity Neighbors

-   Rooks, Bishops, Queen

-   Second-Order (up to X-th Order) Contiguity

    -   Degrees of Seperation

    -   If your second, you have one between

    -   first is basic or right next to you

-   weights matrix - its a matrix 🤠

    -   Can be binary or Distance based (0 for self)

Row Standardisation (or column)

-   This is just to see how many neighbours each thing has

that one graph that uses row standardisation and lagged variables to show how the east is richer than the west. (the sum one includes itself)

## In-Class

Install the GWmodel tool in the dropdown

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr, GWmodel)
```

```{r}
#| eval: false
# import shapefile and parse it to a sf polygon feature object
hunan <- st_read(dsn = "data/data_5/data/geospatial", 
                 layer = "Hunan")
# you should be finding out the projections and doing it accordingly, we didn't because lazy
```

```{r}
#| eval: false
# import csv and parse it into tibble df 
hunan2012 <- read_csv("data/data_5/data/aspatial/Hunan_2012.csv")
# Join Hunan and Hunan_2012 csv 
hunan <- left_join(hunan,hunan2012) #%>%
#  select(1:4, 7, 15)
 # select(1:3, 7, 15, 16, 31, 32)
# this is called data tidying, its so we dont have useless cols
```

```{r}
#| eval: false
write_rds(hunan, "data/data_5/data/hunan.rds")
```

```{r}
#| echo: false
hunan_sf <- read_rds("data/data_5/data/hunan.rds")
```

RDS are used to save the data thats been wrangled, it speeds the process up for subsequent use and uses less memory.

Convert to Spatial Polygon Data Frame

-   the thing we are using doesnt like sf 😭

```{r}
hunan_sp <- hunan_sf %>%
  as_Spatial()
```

```{r}
bw_AIC <- bw.gwr(GDPPC ~ 1,
                 data = hunan_sp,
                 approach = "AIC",
                 adaptive = TRUE,
                 kernel = "bisquare",
                 longlat = T)
```

```{r}
bw_CV <- bw.gwr(GDPPC ~ 1,
                 data = hunan_sp,
                 approach = "CV",
                 adaptive = TRUE,
                 kernel = "bisquare",
                 longlat = T)
```

You can used fixed bandwidth (change adaptive to false)

```{r}
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = bw_AIC,
               kernel = "bisquare",
               adaptive = TRUE,
               longlat = T)
```

Put the good stuff in a dataframe

```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```

Add it to hunan_sf, then you have it in your main data! - will be helpful for graph

```{r}
hunan_gstat <- cbind(hunan_sf, gwstat_df)
# this is appending 
```

Code Displaying it

```{r}
tm_shape(hunan_gstat) +
  tm_fill("GDPPC_LM",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Distribution of geographically weighted mean", 
            main.title.position = "center", 
            main.title.size = 0.5, 
            legend.text.size = 1.2, 
            frame = TRUE)
```
