---
title: "In-class 6"
author: "Lucas Vial"
date: "Sep 23 2024"
date-modified: "Date"
execute: 
  eval: true
  echo: true
  freeze: true
  warning: false
---

# [**Lesson 6: Global and Local Measures of Spatial Autocorrelation**](https://is415-ay2024-25t1.netlify.app/outline/lesson06_outline)

Basically just checking if distributions are random, whether they are considered outliers.

Near things are more related than distant things - spatital dependence and spatial autocorrelation

**Spatial Dependency** - Relationship of variables, statistical dependence

**Spatial Autocorrelation** - Term used to describe the presence of systematic spatial variation, used for clustering pm. Can be across continuous surface, fixed sites or areas.

-   Positive Spatial Autocorrelation: There is a correlation, neighbors are similar. Usually apparent in clustering (Correlation Co-Efficient)

-   Negative Spatial Autocorrelation: Checkerbox patterns, "opposite" of clustering, neighbours are all dissimiliar, More dissimilar than a random output.

## Moran I

Moran I, Z-Value. Describe how features differ from the values in the study area as a whole.

\| + - Clustered \| - - Dispersed \| 0 - Random \|

## Geary C

Geary C, Z-Value. Describes how features differ from their immediate neighbours.

\| \>1 - Dispersed \| \<1 - Clustered \| 1 - Random \|

## Z-Score and P-Values

Alpha value of 0.05, Null Hypothesis, lots of rambling about data timeliness and confidence levels/intervals.

## Getis-Ord Global G

Overall concentration or lack of concentration in all pairs that are neighbours. The variable only takes postitive values. Cluster Detection.

## Local Spatial Autocorrelation Statistics

Collection of geospatial statistical analysis methods to find location related tendency

### LISA - Local Indicator of Spatial Association

Given a set of features and analysis field, the spatial statistics identify spatial clusters of features with high or low values. It also identifies spatial outliers.

High-High \| High-Low \| Low-High \| Low-Low

### Morans Local I

Do this for the first step, its Moran's I, but local. You can then find their p-value to check if the correlation is statistically significant.

Hot and Cold Spots - High Clusters and Low Clusters

Only use the significant ones.

**Break**

## Course Announcements

Take-Home Task 2 is due next week Sunday

-   Theres still more slides at the end to do :(

-   Its based off thailand

-   **No Hands-On Excersise this week or next week**

# In-Class

[sfdep](https://sfdep.josiahparry.com/) package: list columns to make an interface

```{r}
pacman::p_load(sfdep, tidyverse, sf, tmap)
```

```{r}
hunan <- st_read(dsn = "data/6data/data/geospatial", 
                 layer = "Hunan")
hunan2012 <- read_csv("data/6data/data/aspatial/Hunan_2012.csv")
hunan_GDP <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)
```

**Deriving Queens contiguity weights with sfdep methods**

`st_weights()` has three arguments

-   nb: neighbour list objects

-   style: W for row standardized weights, C globally standardised, U is C / n, S is variance-stabilised

-   allow_zero: if true, assigns zero as lagged value to zone without neighbours

```{r}
wm_q <- hunan_GDP %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb,
                         style="W"),
         .before = 1)
#.before puts the column at the front, not the back :) 
```

This adds a column with the neighbors for each row. c(2,3,4) = row 1 is n with 2,3,4

### Computing Global Moran I

Calculate the Global Moran I, the output is a tibble data.frame. You don't normally do this in examples, more of a test :)

```{r}
moranI <- global_moran(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt)
glimpse(moranI)
```

K is average neighbors found, I is the thing ðŸ˜

**Performing Global Moran's I Test**

This is the thing you want to run, instead of computing the stats, you want to perform a test with them. use `global_moran_test` as shown below.

```{r}
global_moran_test(wm_q$GDPPC, wm_q$nb, wm_q$wt)
```

\^ Do analysis stuff's on this

-   Make sure your p-value is low enough (smaller than confidence level at 0.05)

-   THEN use the Moran I statistic to draw conclusions

**Permutation Test**

Monte Carlo should be used to run the simulation. Its supported by sfdep :). You can basically start from this step.

```{r}
# Set a seed for reproducibility
set.seed(1234)
```

Very Similar to the other test but using `global_moran_perm` and adding the nsim, which is set to 99 for how many times its going to be run through.

```{r}
global_moran_perm(wm_q$GDPPC, 
                  wm_q$nb, 
                  wm_q$wt, 
                  nsim = 99)
```

Moran I, statistic is roughly the same, which proves that its stable. We know this because we ran it 100 times ðŸ˜

### Computing Local Moran's I

Here we compute the local Moran's I of GDPPC at county level by using the `local_moran` sfdep package.

```{r}
lisa <- wm_q %>% 
  mutate(local_moran = local_moran(
    GDPPC, nb, wt, nsim = 99),
        .before = 1) %>%
  unnest(local_moran)
```

This creates the lisa data.frame, with lots of helpful statistics, including what region is categorized as (HIGH/HIGH, p, var.....)

-   Lots of p values in this df, It's reccomended that you stick with the same one. (Use p_ii_sim, this will make it the same as your simulation)

-   With the HIGH/LOW columns, the mean and media ones are the important ones.

-   Skewness can help you decide - high skew means you pick median

### **Visualising local Morans I**

```{r}
tmap_mode("plot")
map1 <- tm_shape(lisa)+
  tm_fill("ii")+
  tm_borders(alpha = 0.5)+
  tm_layout(main.title = "Local Moran I")
  # tm_view(set.zoom.limits = c(6,8))
  # tm_layout 
    #main.title 
    # main.title.size 

# This needs to be assigned labels 
map2 <- tm_shape(lisa) +
  tm_fill("p_ii_sim")+ #,
    #      breaks =  This is where you set your sign level legend 
      #      label) +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Significance Level")

tmap_arrange(map1, map2, ncol = 2)
```

### Lisa Map Visualization

The LISA Map Visualization is a combination of the p significance level and local moran I, it will graph the Moran correlations when the significance meets its required level (null hypothesis)

```{r}
#| warning: false
#| error: false
lisa_sig <- lisa %>% 
  filter(p_ii < 0.05)
tm_shape(lisa)+
  tm_polygons()+
  tm_borders(alpha=0.5)+
tm_shape(lisa_sig) +
  tm_fill("mean")+
  tm_borders(alpha=0.4)+
  tm_layout(main.title = "LISA map")
```

## Hot and Cold Spot Analysis

Computing Local G\* statistics (One with and without asterix)

This is in the same format as the Local Moran I statistics

```{r}
wm_idw <- hunan_GDP %>%
  mutate(nb = st_contiguity(geometry),
         wts = st_inverse_distance(nb, geometry,
                                   scale = 1,
                                   alpha = 1),
         .before = 1)
```

Similar process to LISA/Moran I stat creation

```{r}
set.seed(1234)
HCSA <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    GDPPC, nb, wt, nsim = 99),
    .before = 1) %>%
  unnest(local_Gi)
HCSA
```

**Visualisation**

```{r}
tm_shape(HCSA)+
  tm_fill("gi_star")+
  tm_borders(alpha=0.5)+
  tm_layout(main.title = "G*")
```

```{r}
HCSA_sig <- HCSA %>% 
  filter(p_sim < 0.05)

tm_shape(HCSA)+
  tm_polygons()+
  tm_borders(alpha=0.5)+
tm_shape(HCSA_sig)+
  tm_fill("gi_star")+
  tm_borders(alpha=0.4)
```
