---
title: "In-class 7"
author: "Lucas Vial"
date: "Sep 30 2024"
date-modified: "Date"
execute: 
  eval: false
  echo: true
  freeze: true
  warning: false
---

# Web-Enabled Geospatial Analytics with R Shiny

Automated Data Analysis

Rich Internet Applications (RIA)

Shiny is the thing that makes your R code, turn to web-based

-   Server Instructions (R)

-   User Interface (HTML)

```{r}
library(shiny)
runExample("01_hello")
```

Basic Shiny Skeleton

```{r}
library(shiny)

ui <- fluidPage()

# Define server logic required to draw a histogram
server <- function(input, output) {}

# Run the application 
shinyApp(ui = ui, server = server)

```

Connecting to Shinyapps.io

```{r}
install.packages('rsconnect')
```

```{r}
rsconnect::setAccountInfo(name='lucas-v', token='E44A25795BA5772DE670A9A12F3E60AB', secret='CmGPPaQGtlSmFkHoh8/NFUVG0JYjGRIVlz2UEwhN')

```

```{r}
library(rsconnect)
    rsconnect::deployApp('shiny/GAShiny')
```

Ok, I wasn't meant to do it like this - this is how it says to do it on the website, other way is rendering and hitting publish

<https://lucas-v.shinyapps.io/gashiny/>
