---
title: "Take Home 3"
author: "Lucas Vial"
date: "Nov 2 2024"
date-modified: "Date"
execute: 
  eval: false
  echo: true
  freeze: true
  warning: false
---

### Task Goals

Import R Packages

Prepare and Test R code so output is correct

Determine and prepare parameters and outputs for shiny

select appropriate shiny components

### Shiny Imports

```{r}
install.packages('rsconnect')
```

```{r}
rsconnect::setAccountInfo(name='lucas-v', token='E44A25795BA5772DE670A9A12F3E60AB', secret='CmGPPaQGtlSmFkHoh8/NFUVG0JYjGRIVlz2UEwhN')
```

```{r}
library(rsconnect)
    rsconnect::deployApp('shiny/takehome3')
```

```{r}
pacman::p_load(sfdep, tidyverse, sf, tmap)

```

```{r}
adm1 <- readRDS("shiny/takehome3/data/papua_adm1.rds")
adm2 <- readRDS("shiny/takehome3/data/papua_adm2.rds")
ps <- readRDS("shiny/takehome3/data/points_study.rds")
```

```{r}
library(shiny)
runApp("shiny/takehome3")
```

### UI Design

Its the same as the proposed design, Data ranges, bandwidth selection method, smooting kernel function, second-order analysis method, number of simulations and edge correction for monte carlo. The design has 2 visualizations with the KDE map and the monte carlo simulation.

[https://lucas-v.shinyapps.io/takehome3](https://lucas-v.shinyapps.io/takehome3/)/
